Tropical Mathematics and the Lambda Calculus

### 2023-01-24 section2.tex
commit

These observations have led to the study of linear $\lambda$-calculi with \emph{graded} exponentials, like the system $\mathsf{Fuzz}$ \cite{Reed2010}, inspired from Girard's Bounded Linear Logic \cite{Girard92tcs}, which have been applied to the study of differential privacy \cite{Gaboardi2013, Gaboardi2017}. The types of such systems are defined by combining linear constructors with a \emph{graded exponential comonad} $!_{r}(-)$ \cite{Katsumata2018}.
In the following sections we will sometimes make reference to a basic graded type system, that we call $\BSTLC$, for bounded higher-order programs, with types defined via $A::= o  \mid   !_{n}A\multimap A  $, 
where $n\in \BB N$. Intuitively, $!_{ n}A\multimap B$ is the type of functions from $A$ to $B$ that may use their input \emph{at most} $n$ times. More details about $\BSTLC$ are provided in the Appendix.

@@

Yet, what if these polynomials were tropical ones, i.e.~piecewise linear functions? This way, \eqref{eq:taylor} could really be interpreted as decomposition of $\lambda$-terms via limits (indeed, an $\inf$s) of Lipschitz maps. In other words, unbounded term application could be seen 
as a limit of \emph{more and more sensitive} operations.

>>>>

These observations have led to the study of linear $\lambda$-calculi with \emph{graded} exponentials, like $\mathsf{Fuzz}$ \cite{Reed2010}, inspired from Girard's Bounded Linear Logic \cite{Girard92tcs}, which have been applied to the study of differential privacy \cite{Gaboardi2013, Gaboardi2017}. The types of such systems are defined by combining linear constructors with a \emph{graded linear exponential comonad} $!_{r}(-)$ \cite{Katsumata2018}.
In the following sections we will sometimes make reference to a basic graded type system, that we call $\BSTLC$, for bounded higher-order programs, with types defined via $A::= o  \mid   !_{n}A\multimap A  $, 
where $n\in \BB N$. Intuitively, $!_{ n}A\multimap B$ is the type of functions from $A$ to $B$ that may use their input \emph{at most} $n$ times. More details about $\BSTLC$ are provided in the Appendix.

@@

Yet, what if these polynomials were tropical ones, i.e.~piecewise linear functions? This way, \eqref{eq:taylor} could really be interpreted as a decomposition of $\lambda$-terms via limits (indeed, $\inf$s) of Lipschitz maps. In other words, unbounded term application could be seen 
as a limit of \emph{more and more sensitive} operations.

### 2023-01-24 section4.tex
commit

%many properties of these functions still have to be well-understood. 
In this section we establish some tropological and metric properties of tropical Laurent series, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
The literature on power series in tropical mathematics is very recent (see e.g.~\cite{Porzio2021}) and several results we prove in this section are, to our knowledge, new.

>>>>

%many properties of these functions still have to be well-understood. 
In this section we establish some topological and metric properties of tropical Laurent series, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
The literature on power series in tropical mathematics is very recent (see e.g.~\cite{Porzio2021}) and several results we prove in this section are, to our knowledge, new.
